On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-04 main.tex
minor
intro

%%
%% This is file `sample-xelatex.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-xelatex.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}

% packages 
%\usepackage[options ]{algorithm2e}
\usepackage[ruled,vlined]{algorithm2e}
%\usepackage{algorithm,algpseudocode,float}
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage{enumitem}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\bz{\textit{\textbf{0}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\pb{\textit{\textbf{p}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\pr{\mathbb{P}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textrm{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\kt{\widetilde{k}}
\def\dt{s}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.

@@

\author{Eric Schost}
\affiliation{%
  \institution{Cheriton School of Computer Science}
  \city{University of Waterloo}
}
\email{eschost@uwaterloo.ca}
%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Elliott, Giesbrecht, and Schost.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  We present a full analysis of an efficient algorithm for the
  computation of at least one point in each connected component of a
  smooth real hypersurface, taking into account bit complexity and
  all coefficient growth.  This problem is a basic and important
  subroutine used in semi-algebraic geometry. For instance, it is used
  in determining an upper bound on the number of connected components
  of a real hypersurface. The algorithm requires some genericity
  properties, which we ensure through random changes of variables that
  produce generic coordinate systems. In order to ensure these
  properties with arbitrarily high probability, we developed a new
  quantitative extension of Thom's weak transversality theorem, and a
  new quantitative Noether normalization statement for polar
  varieties. These statements enable us to give a precise probability
  analysis of a monte carlo algorithm, where we can guarantee
  correctness with arbitrarily high probability.

  Algorithms for deciding connectivity queries in real algebraic
  geometry have been developed in a series of papers
  \cite{EMP,SchostMohabBabySteps2011,SchostMohabBabySteps2014,TWT}. However, these algorithms are based in the
  arithmetic complexity model, and analyses of the bit complexity
  questions were left for future work.  Our results are an important
  first step in this analysis.

\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

%\ccsdesc[500]{Computer systems organization~Embedded systems}
%\ccsdesc[300]{Computer systems organization~Redundancy}
%\ccsdesc{Computer systems organization~Robotics}
%\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Real algebraic geometry, transversality, Noether normalization, bit complexity}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.

@@

@@

\subsection{Problem statement}
Let $f \in \mathbb{Q}[X_1,\hdots,X_n]$ be squarefree with total degree $D$ and $V(f)$ smooth. We provide bit-complexity estimates for computing one point in each connected component of $V(f)\cap \R^n.$ Its boolean complexity is $\hdots$.
\subsection{Motivation}
\subsubsection{Applications}
Computing one point in each connected component of the real part of a hypersurface is a basic and important subroutine used in semi-algebraic geometry. For instance, it is used in determining an upper bound on the number of connected components of a real hypersurface, and it is used in determining whether or not a complex hypersurface has real solutions. 

@@

Algorithms for deciding connectivity queries in real algebraic geometry have been developed in a series of papers \cite{EMP,SchostMohabBabySteps2011,SchostMohabBabySteps2014,TWT}. However, these algorithms are based in the arithmetic complexity model, and bit complexity questions have been left for future work. As bit complexity provides a far more realistic model of computation compared to arithmetic complexity,  which only counts each operation at unit cost, our results therefore begin to fill an important missing gap in the research. 
\subsubsection{Quantitative genericity statements}
\subsection{Main result}
Suppose that $f\in\Q[X_1\hdots,X_n]$ is squarefree and satisfies $\deg f \leq D, \htt(f) \leq s,$ and assume that $f$ is given by a straight-line program $\Gamma$ of size $L$ with integer constants of height at most $b.$ There exists a randomized algorithm that takes $\Gamma, d,$ and $s$ as input and produces a zero-dimensional parameterization of a set that contains at least one point in each connected component of $V(f) \cap \R^n$ with probability at least $1/2$. Otherwise the algorithm either produces a subset of the points or FAIL. In any case, the algorithm uses 
O^{\sim}(kn(Lb+D^{2n}(s+D)(L+n^3D+n^4)))
boolean operations, with $k \geq -\lg(1-2^{(-1/n)})$. The polynomials in the output have degree at most $D^n,$ and height at most
O^{\sim}((s+D)D^n + nD^n).
The algorithm can error on one side by producing a proper subset of the points. Running the algorithm $\kt$ times gives a list of $n$ zero-dimensional parameterizations, among which the highest cardinality output includes at least one point in each connected component with probability at least $1-(1/2)^{\kt}.$
\section{Notation and preliminaries}
\subsection{Algebraic sets}

@@

\subsubsection{Dimension and degree}

@@

\subsection{Bit-size, height and degree}

@@

For $r \leq n,$ $\textit{\textbf{X}}_{\leq r}$ denotes $[X_1,\hdots,X_r]$ and $\textit{\textbf{X}}$ denotes $[X_1,\hdots,X_n]$. 
For $a=\frac{u}{v}$ in $\mathbb{Q}-\{0\},$ the  \textit{height} of $a,$ $\htt(a),$ is the maximum of $\log(|u|)$ and $\log(|v|),$ where $u \in \mathbb{Z}$ and $v \in \mathbb{N}$ are coprime. If $v$ is the minimal common denominator of all non zero coefficients of $f$, then the \textit{height} of $f, \htt(f),$ is defined as the maximum of the logarithms of $v$ and of the absolute values of the coefficients of $vf$. 
We write $\htt(\textit{\textbf{f}})=(\htt(f_1),\hdots,\htt(f_N))$ and assume that $\htt(\textit{\textbf{f}}) \leq \textit{\textbf{s}} = (S_1,\hdots,s_N),$  with $\htt(f_i) \leq s_i$ for all $i \in \{ 1,\hdots,N\}.$
\subsection{Data structures}
\subsubsection{Strait-line programs}
\subsubsection{Zero-dimensional parameterizations}

@@

@@

Consider $f \in \mathbb{Q}[X_1,\hdots,X_n]$ with total degree $D$.  

@@

An equidimensional algebraic set $X$ of dimension $d$ is in \textit{Noether position} for $\pi_d$ when the extension 
\C[X_1,\hdots,X_d] \rightarrow \C[X_1,\hdots,X_n]/I(V)
is injective and integral.  Then, for any $\xb \in \C^d,$ the fiber 
$V \cap \pi_d^{-1}(\xb)$ has dimension zero. For $i\in\{1,\hdots,n\}$, and for a randomly chosen point $p_{i-1} = (x_1,\hdots,x_{i-1}) \in \Q^{i-1}$, we say that $f$ satisfies $\textbf{H}_i^{'}$ if 

@@

\item $\langle X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}\rangle $ is radical; 
\item $V(X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n})$ is finite;
\item $V(X_1-x_1,\hdots,X_{i-1}-x_{i-1}, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n})$ is smooth;

@@

Suppose that $f \in \Q[X_1,\hdots,X_n]$ satisfies \textbf{H}. Then there exists a Zariski closed set $\D \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{\td}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\Delta$ ( $\Q^{n^2}-\Delta$?) then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$

@@

Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{\td}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$

@@

It will sometimes be useful for us use in replace of the matrix $A$, a matrix representing indeterminants: let $\A$ be an $n \times n$ matrix of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in \mathbb{Q}(\A_{i,j})[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$ It follows from Proposition 5.1, that for a generic $A \in \C^{n^2},$ $0$ is a regular value of $\Phi_{A}$.  The proof of \cite[Theorem B.3]{TWT}, which establishes the existence of $\Delta \subset \C^{n^2},$ applies \textit{Sard's lemma}, which shows that the critical values of a polynomial mapping are contained in a hypersurface. It follows from Sard's lemma that, for the matrix of indeterminants $\A,$ $0$ is a regular value of $\Phi_{\A}.$ Otherwise it could not be that $0$ is a regular value of $\Phi_A,$ for a generic $A\in \C^{n^2}.$ The proof is routine and we omit details. It then follows that 

@@

\jac_{\xb}(\Phi_{\A}) = \jac_{\xb}\left(f^{\A}(\xb),\frac{\pa f^{\A}(\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f^{\A}(\xb)}{\pa X_n}\right)

@@

has full rank for all $\xb \in V(f^{\A})$, and thus $f^{\A}$satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$ by the Jacobian criterion \cite[Theorem 16.19]{ECA}. We will rely on this fact in Section 8.

@@

We assume our input polynomial $f$ satisfies \textbf{H} so that $f$ is squarefree with $V(f) \cap \mathbb{R}^n$ smooth. Thus, by Proposition 2.4,  $W(\pi_i,V(f))$ is defined by the vanishing of $f,\frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}.$ If the Jacobian of this set of polynomials has full rank at all $\xb \in V(f) \cap \mathbb{R}^n,$ then it follows by the Jacobian criterion \cite[Theorem 16.19]{ECA} that $f$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$. 

@@

There exists a Zariski closed set $\Delta_1 \subset \C^{n^2}$ of degree at most $D^{2n^2}$ with the property that, if $A \in \C^{n^2} - \Delta_1$ then for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$.

@@

\subsection{Sketch of the proof of Theorem 5.1}

@@

For each $i \in \{1,\hdots,n\}$, let $\Phi_i$ be the polynomial mapping 

@@

\Phi_i : (\xb,A) \mapsto \left(f^A(\xb), \frac{\partial f^A}{\partial X_{i+1}}(\xb),\hdots,\frac{\partial f^A}{\partial X_n}(\xb)\right),

@@

so that $\Phi_{i,A}^{-1}(0)= W(\pi_i,V^A).$ We want to apply Theorem 5.1, and thus we show that $\bz$ is a regular value of $\Phi_i$, where then it follows that $\Xi_i \subset \C^{n^2}$ exists with the property that if $A \in \C^{n^2}-\Xi_i$ then  

@@

\jac_{\xb}(\Phi_{i,A}) = \jac_{\xb}\left(f^A(\xb),\frac{\pa f^A(\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f^A(\xb)}{\pa X_n}\right)

@@

\Delta_1 = \bigcup_{i=1}^n \Xi_i 

@@

where we can bound the degree by observing that 

>>>>

\documentclass[sigconf]{acmart}

% test 
\usepackage{mathrsfs}
\usepackage{bm}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\ZZ{\mathbb{Z}}
\def\R{\mathbb{R}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textrm{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\dt{s}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\scrQ{\ensuremath{\mathscr{Q}}}
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

\usepackage[ruled,vlined]{algorithm2e}
\usepackage{chngcntr}
\usepackage{titlesec}

@@

\author{\'Eric Schost}
\affiliation{%
  \institution{Cheriton School of Computer Science}
  \city{University of Waterloo}
}
\email{eschost@uwaterloo.ca}
%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Elliott, Giesbrecht, and Schost.}

\begin{abstract}
  We present a full analysis of the bit complexity of an efficient
  algorithm for the computation of at least one point in each
  connected component of a smooth real hypersurface. This is a basic
  and important operation in semi-algebraic geometry: it gives an
  upper bound on the number of connected components of a real
  hypersurface, and is also used in many higher level algorithms.

  Our starting point is an algorithm by Safey El Din and Schost ({\em
    Polar varieties and computation of one point in each connected
    component of a smooth real algebraic set}, ISSAC'03). This
  algorithm uses random changes of variables, that are proved to
  generically ensure certain desirable geometric properties. The
  complexity of the algorithm was given in an algebraic complexity
  model; the analysis of the bit complexity and the error probability
  were left for future work.

  Our paper answer these questions. Our main contribution is a
  quantitative analysis of several genericity statements, such as
  Thom's weak transversality theorem or Noether normalization
  properties for polar varieties.
\end{abstract}

\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\setcounter{secnumdepth}{4}
\renewcommand{\theparagraph}{\thesubsection.\arabic{paragraph}}
\counterwithin{paragraph}{subsection} % makes paragraph depend on subsection
\titleformat{\paragraph}[runin]{\normalfont\normalsize\bfseries}{\theparagraph.}{1em}{}
\titlespacing*{\paragraph}{0em}{1ex}{1em}
\newcommand{\pref}[1]{{\bf\ref{#1}}}

\keywords{Real algebraic geometry, transversality, Noether
  normalization, bit complexity}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph*{Background and problem statement.}
Computing one point in each connected component of a real algebraic
set $S$ is a basic subroutine in real alebraic and semi-algebraic
geometry; it is also useful in its own right, since it allows one to
decide if $S$ is empty or not. In this paper, we consider only the
case where $S$ is given as $S=V \cap \R^n$, where $V=V(f) \subset
\C^n$ is a complex hypersurface defined by a squarefree polynomial $f
\in \ZZ[X_1,\dots,X_n]$.

Algorithms for this task have been known for decades, and their
complexity is to some extent well understood. Suppose that $f$ has
degree $d$, and coefficients of bit-size $h$. Without making any
assumption on $f$, the algorithm given
in~\cite[Section~13.1]{BaPoRo03} solves our problem using $d^{O(n)}$
operations in $\Q$; in addition, the output of the algorithm is
represented by polynomials of degree $d^{O(n)}$, with coefficients of
bit-size $hd^{O(n)}$.  The key idea behind this algorithm goes back
to~\cite{GrVo88}: sample points are found through the computation of
critical points of well-chosen functions on $V(f)$.

The number of connected components of $V(f)$ admits the lower bound
$d^{\Omega(n)}$, so up to polynomial factors, this result is
optimal. However, due to the generality of the algorithm, the constant
hidden in the exponent $O(n)$ in its runtime turns out to be rather
large: the algorithm relies on infinitesimal deformations, that affect
runtime non-trivially.

In this paper, we will work under the additional assumption that
$V=V(f)$ is a {\em smooth} complex hypersurface. We place ourselves in
the continuation of the line of work initiated by~\cite{BaGiHeMb97}:
that reference deals with cases where $V$ is smooth and $V \cap \R^n$
is compact, pointing out how {\em polar varieties} (that were
introduced in the 1930's in order to define characteristic
classes~\cite{Piene78,Teissier88}) can play a role in effective real
geometry. This paper was extended in several directions: to $V$ being
a smooth complete intersection, still with $V\cap \R^n$
compact~\cite{BaGiHeMb01}, then without the compactness
assumption~\cite{EMP,BaGiHePa05}; the smoothness assumption was then partly
dropped in~\cite{BaGiHe14,BaGiHeLePa12}.

The starting point in our paper is the algorithm in~\cite{EMP}. In the
hypersurface case, its runtime is $d^{(4+o(1))n}$ operations in $\Q$;
as with many results in this line of work, the algorithm is
randomized: the algorithm needs to assume that we are in generic
coordinates; this is done by applying a random change of coordinates
prior to all computations. In addition, it relies on algorithms for
solving systems of polynomial equations that are themselves
randomized.  Altogether, this procedure chooses $n^{O(1)}$ random
vectors, each of them in an affine space of dimension $n^{O(1)}$;
every time a choice is made, failure corresponds to our chosen vector
lying in a hypersurface of the parameter space.

This paper revisits the algorithm in~\cite{EMP} (in the hypersurface
case) and gives a complete analysis of its probality of success and
its bit complexity.

\paragraph*{Data structures.} As in references such
as~\cite{GiHeMoPa95,GiHaHeMoMoPa97,GiHeMoMoPa98,BaGiHeMb97,EMP}, the
input polynomial $f$ is given by means of a {\em straight-line
  program}, that is, a sequence of operations $+,-,\times$ that
evaluates $f$, on the input variables $X_1,\dots,X_n$; the {\em
  length} of a straight-line program of such an objet is simply the
number of operations it performs. Any polynomial of degree $d$ in $n$
variables can be computed by a straight-line program of length $O\big
({{d+n} \choose n}\big)$ (evaluate all monomials of degree up to $d$
in $n$ variables, multiply them by their respective coefficients and
sum the result). However, some inputs may be given by shorter
straight-line program, and the algorithm is able to benefit from it.

The output of the algorithm is a finite set in $\C^n$ defined over
$\Q$. To represent it, we rely on a widely used data structure based
on univariate
polynomials~\cite{Kronecker82,Macaulay16,GiMo89,GiHeMoPa95,ABRW,GiHaHeMoMoPa97,GiHeMoMoPa98,Rouillier99}.
Consider a zero-dimensional algebraic set $S \subset \C^n$ defined
over $\Q$. A {\em zero-dimensional parametrization}
$\scrQ=((q,v_1,\dots,v_n),\lambda)$ of $S$ consists in polynomials
$(q,v_1,\dots,v_n)$, such that $q\in \Q[T]$ is monic and squarefree,
all $v_i$'s are in $\Q[T]$ and satisfy $\deg(v_i) < \deg(q)$, and in a
$\Q$-linear form $\lambda$ in variables $X_1,\dots,X_n$, such that
\begin{itemize}
\item $\lambda(v_1,\dots,v_N)=T q' \bmod q$, where $q'=\frac{\partial
  q}{\partial T}$;
\item we have the equality
  $S=\left \{\left(
      \frac{v_1(\tau)}{q'(\tau)},\dots,\frac{v_N(\tau)}{q'(\tau)}\right
    ) \ \mid \ q(\tau)=0 \right \}.$
\end{itemize}
The constraint on $\lambda$ says that the roots of $q$ are the values
taken by $\lambda$ on $S$. The parametrization of the coordinates by
rational functions having $q'$ as a denominator goes back
to~\cite{Kronecker82,Macaulay16}: as pointed out in~\cite{ABRW}, it
allows one to control precisely the size of the coefficients of
$v_1,\dots,v_n$.

\paragraph*{Main result.} To state our main result, we need to define 
the {\em height} of rationals, and polynomials with rational
coefficients.

The {\em height} of a non-zero rational $a=u/v$ is the maximum of
$\ln(|u|)$ and $\ln(|v|),$ where $u \in \mathbb{Z}$ and $v \in
\mathbb{N}$ are coprime. For a polynomial $f$ with rational
coefficients, if $v \in \mathbb N$ is the minimal common denominator
of all non-zero coefficients of $f$, then the \textit{height}
$\htt(f)$ of $f$ is defined as the maximum of the logarithms of $v$
and of the absolute values of the coefficients of $vf$.

\begin{theorem}
  Suppose that $f\in\ZZ[X_1\hdots,X_n]$ is squarefree, of degree
  at most $d$ and height at most $s$; assume that $f$ is given by a
  straight-line program $\Gamma$ of size $L$ with integer constants of
  height at most $b.$ There exists a randomized algorithm that takes
  $\Gamma, d,$ and $s$ as input and produces a zero-dimensional
  parameterization of a set that contains at least one point in each
  connected component of $V(f) \cap \R^n$ with probability at least
  $\hdots$. Otherwise the algorithm either produces a subset of the
  critical points or FAIL. In any case, the algorithm uses
  \[
  \hdots  
  \]
  boolean operations.
\end{theorem}

The algorithm is Monte Carlo and can error on one side by producing a
proper subset of the points. Running the algorithm $k$ times gives a
list of outputs among which the highest cardinality set includes at
least one point in each connected component with probability at least
$\hdots.$

%% Let $f \in \mathbb{Q}[X_1,\hdots,X_n]$ be squarefree with total degree
%% $d$, and such that $V(f)$ smooth. We provide bit-complexity estimates
%% for computing one point in each connected component of $V(f)\cap
%% \R^n.$ Its boolean complexity is $\hdots$.

@@

Algorithms for deciding connectivity queries in real algebraic
geometry have been developed in a series of papers
\cite{a,b,c,d}. However, these algorithms are based in the arithmetic
complexity model, and bit complexity questions have been left for
future work. As bit complexity provides a far more realistic model of
computation compared to arithmetic complexity, which only counts each
operation at unit cost, our results therefore begin to fill an
important missing gap in the research.

%
%
%

@@

@@

@@

@@

\section{The Main Algorithm}
\begin{algorithm}
\nl Choose $S \subset \mathbb{Q}$ and $A \in S^{n^2}$ with $|S|\geq $?\;
\nl Use \cite{AD} to compute the partial derivatives $\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}$\;
Apply the symbolic homotopy algorithm from \cite{SH} with input $\{X_1,\hdots,X_i,f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n} \}$ and $D$ to obtain a zero dimensional parameterization of $W(\pi_i,V^A) \cap \pi_{i-1}^{-1}(0)$.;
\nl \Return $\cup_{i=1}^n W(\pi_i,V^A) \cap \pi_{i-1}^{-1}(0)$;

@@

Consider $f \in \mathbb{Q}[X_1,\hdots,X_n]$ squarefree with total degree $D$ and $V(f)$ smooth. We say that $V$ is in \textit{Noether position} for $\pi_D$ when the extension 
\C[X_1,\hdots,X_D] \rightarrow \C[X_1,\hdots,X_n]/I(V)
is injective and integral.  Then, for any $\xb \in \C^D,$ the fiber 
$V \cap \pi_D^{-1}(\xb)$ has dimension zero. 

@@

For $i\in\{1,\hdots,n\}$, we say that $f$ satisfies $\textbf{H}_i^{'}$ if 

@@

\item $\langle X_1,\hdots,X_i, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}\rangle $ is radical; 
\item $V(X_1,\hdots,X_i, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n})$ has dimension $0$;
\item $V(X_1,\hdots,X_i, f,  \frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n})$ is smooth;

@@

Suppose that $f$ satisfies \textbf{H}. Then there exists a Zariski closed set $\frak{X} \subset \mathbb{C}^{n^2}$ of degree at most $D^{\textrm{?}}$ such that, if $\textbf{A}$ is chosen from  $\C^{n^2}-\frak{X}$ then $f^A$ satisfies $\textbf{H}_i^{'}$ for all $i \in \{1,\hdots,n\}.$

@@

Fix $S \subset \mathbb{Q}$ with $|S| \geq \epsilon^{-1} D^{\textrm{?}}$ and $\epsilon > 0$. Then for $A\in S^{n^2}$ chosen randomly, the probability that $f^A$ satisfies $H_i^{'}$ for all $i \in \{1,\hdots,n\}$ is at least $1-\epsilon.$

@@

It will sometimes be useful for us to let the matrix $A$ represent new
indeterminates: let $\A$ be an $n \times n$ matrix of new
indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in
\mathbb{Q}(\A_{i,j})[\textbf{X}]$ as $f(\A\textbf{X})$ and
$V^\A=V(f^\A).$ It follows from Proposition 4.2 that for a generic $A
\in \C^{n^2},$ $0$ is a regular value of $\Phi_{A}$.  The proof of
\cite[Theorem B.3]{NO}, which shows the existence of $\Delta \subset
\C^{n^2},$ applies \textit{Sard's lemma}, which shows that the
critical values of a polynomial mapping are contained in a
hypersurface. It follows from Sard's lemma that for the matrix of
indeterminants $\A,$ $0$ is a regular value of $\Phi_{\A}.$ Otherwise
it could not be that $0$ is a regular value of $\Phi_A,$ for a generic
$A\in \C^{n^2}.$ Therefore,

@@

\jac_{\xb}(\Phi_{\A}) = \jac_{\xb}\left(f^{\A}(x),\frac{\pa f^{\A}(x)}{\pa X_{i+1}},\hdots,\frac{\pa f^{\A}(x)}{\pa X_n}\right)

@@

has full rank for all $\xb \in V(f^{\A})$, and thus $f^{\A}$satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$. We will rely on this fact in Section 5. 

@@

We assume our input polynomial $f$ satisfies \textbf{H} so that $f$ is squarefree with $V(f) \cap \mathbb{R}^n$ smooth. Thus, by Proposition 2.4,  $W(\pi_i,V(f))$ is defined by the vanishing of \[\fb := \left(f,\frac{\pa f}{\pa X_{i+1}},\hdots,\frac{\pa f}{\pa X_n}\right).\]
If the Jacobian of $\fb$ has full rank at all $x \in V(f) \cap \mathbb{R}^n,$ then it follows by the Jacobian criterion [4, Theorem 16.19] that $f$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$. 

@@

There exists a Zariski closed set $\Delta \subset \C^{n^2}$ of degree at most $D^n$ with the property that, if $A \in \C^{n^2} - \Delta$ then for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$.

@@

@@

We will let $\Phi$ be the mapping 

@@

(\xb,A) \mapsto \left(f^A(\xb), \frac{\partial f^A}{\partial X_{i+1}}(\xb),\hdots,\frac{\partial f^A}{\partial X_n}(\xb)\right),

@@

so that 

@@

\Phi_A^{-1}(0)= W(\pi_i,V^A).

@@

\jac_{\xb}(\Phi_A) = \jac_{\xb}\left(f^A(x),\frac{\pa f^A(x)}{\pa X_{i+1}},\hdots,\frac{\pa f^A(x)}{\pa X_n}\right)

@@

has full rank for all $\xb \in V(f^A)$. Then, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i'(1),\textbf{H}_i'(2)$ and $\textbf{H}_i'(3)$.
